---
title: "Homework 10"
author: "Shahriar Shohid Choudhury"
date: "January 14, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

For all exercises, use the KiGGS dataset.

#kiGGS dataset

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
```

## Exercise 1: Logistic regression

Choose 1 suitable outcome variable of interest and 3 predictors, and compute a logistic regression model. Interpret the results: which predictor is associated with the outcome and what is the strength of association (odds ratio)? Also, is the model a good fit i.e. can the outcome be predicted well (look at the misclassification table for this)?

```{r}
# Variable preparation for logistic regression
str(dat$Maanzahl)
table(dat$Maanzahl)
str(dat$sex)
str(dat$age2)
str(dat$STALA)

dat$measles <- dat$Maanzahl > 0

#Factorization
dat$sex   <- factor(dat$sex, labels = c("boys", "girls"))
dat$age2  <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
dat$STALA <- factor(dat$STALA, labels = c("rural", "small_city", "medium_city", "urban"))
```

```{r}
# logistic regression:
fit1 <- glm(measles ~ sex + STALA + age2, family = binomial(link = "logit"), data = dat)
summary(fit1)
library(jtools)
jtools::summ(fit1, exp = T, confint = T, model.fit = F, digits = 3)
```
```{r}
# Computing the odds ratio:
exp(coef(fit1))
```
```{r}
# model fit:
pred1 <- predict(fit1, type = "response")
summary(pred1)

measles.p.pred1 <- rep(NA, length(dat$measles))
measles.p.pred1[as.numeric(names(pred1))] <- pred1
measles.pred1 <- measles.p.pred1 > mean(measles.p.pred1, na.rm = TRUE)

table(measles.pred1, dat$measles)

#The odds ratios in small and medium cities for measles vaccinations are 1.048 and 1.1, respectively. However, the ratio in urban areas is substantially lower than in the other two, at 0.866. When compared to other age groups, the 14 to 15 year old group has the greatest ratio. This model misclassified a significant portion of the samples, as seen by the missclassification table. We can conclude that the model neither fits the data well nor does it anticipate the result well.

#I am a bit confused about this question. hence i will clear my thoughts in the next tutorial.
```

## Exercise 2: Poisson regression

Predict the amount of measles vaccinations (Maanzahl) by the number of siblings (e006B1), the sex and age of the children (sex, age2), place of residence (STALA, OW) and the monthly household income (e093), using a Poisson regression model. Interpret the results. Which variables are associated with the outcome? Is the model a good fit to the data?

```{r}
#Preparation - check variables
str(dat$Maanzahl)
str(dat$e006B1)
str(dat$sex)
str(dat$age2)
str(dat$STALA)
str(dat$OW)
str(dat$e093)
```
```{r}
# Format variables
dat$Maanzahl <- as.numeric(as.character(dat$Maanzahl))
dat$e006B1<- factor(dat$e006B1, labels = c("no sibling", "1 sibling", "2 or more siblings"))
dat$sex   <- factor(dat$sex, labels = c("boys", "girls"))
dat$age2  <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
dat$OW    <- factor(dat$OW, labels = c("east", "west"))
dat$STALA <- factor(dat$STALA, labels = c("rural", "small_city", "medium_city", "urban"))
```

```{r}
# Look at variables
table(dat$Maanzahl)
table(dat$e006B1)
table(dat$sex)
table(dat$age2)
table(dat$STALA)
table(dat$OW)
table(dat$e093)
```
```{r}
# look at distribution of outcome
hist(dat$Maanzahl)
mean(dat$Maanzahl, na.rm = TRUE)
var(dat$Maanzahl, na.rm = TRUE)

# mean and var are not very different. Mean value is 1.632298 and var is 0.6555274 so (mean>var, below we will fit a Poisson regression model for illustration)
```
```{r}
# Poisson Regression: first predict dat$Maanzahl only by sex:
fit1 <- glm(dat$Maanzahl ~ sex, family = poisson(link = "log"), data = dat)

# results:
summary(fit1)

# for interpretation: e^ to have interpretation on the scale of frequencies (and not of log-frequencies):

exp(coef(fit1))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit1, exp = T, confint = T, model.fit = F, digits = 3)
```
```{r}
# Predicted frequencies by the model:
pred_1 <- predict(fit1, type="response")

# Compare these with the observed frequencies:
predictions_1 <- data.frame(obs = dat$Maanzahl[1:17640 %in% names(pred_1)], pred_1 = pred_1)
cor(predictions_1$obs, predictions_1$pred_1)
plot(predictions_1$obs, predictions_1$pred_1)

# The model doesn't seem to predict anything - sex is only very weakly associated with the amount of measles vaccinations
```
```{r}
# fit model with all variables:
fit2 <- glm(dat$Maanzahl ~ e006B1 + sex + age2 + STALA + OW + e093, family = poisson(link = "log"), data = dat)

# results:
summary(fit2)
exp(coef(fit2))
library(jtools)
jtools::summ(fit2, exp = T, confint = T, model.fit = F, digits = 3)
```
```{r}

# Predictions:
predictions_2 <- data.frame(emp = dat$Maanzahl[1:17640 %in% names(predict(fit2, type="response"))], exp = predict(fit2, type="response"))
cor(predictions_2$emp, predictions_2$exp)
plot(predictions_2$emp, predictions_2$exp)

#The results appear to be weakly associated with sex however strongly correlated with e006B1, age2, and OW. The misclassification table shows that there is a correlation of 0.547 between the predicted and actual values, which is sufficient evidence that the model does not well match the data and is unable to predict the outcome. So we can say that its neither a perfect model nor the bad one.
```


## Exercise 3: Negative Binomial regression (optional)

Predict the amount of measles vaccinations (Maanzahl) by the number of siblings (e006B1), the sex and age of the children (sex, age2), place of residence (STALA, OW) and the monthly household income (e093), using a Negative Binomial regression model. Interpret the results. Which variables are associated with the outcome? Is the model a good fit to the data?
