---
title: "Homework 11"
author: "Shahriar Shohid Choudhury"
date: "January 22, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

For all exercises, use the KiGGS dataset.

## Exercise 1: Analysis of variance

Choose serum glucose levels (GLUCX) as outcome variable of interest, the factors BMI group (bmiKH) and sex (sex), and physical wellbeing (kw100.e) as quantitative covariate (=metric variable), and compute 5 ANOVA models: (i-ii) one model with each factors separately, (iii) one model with both factors, (iv) one model with both factors and their interaction, (v) one model with both factors, their interaction, and the covariate.

Choose at least one aspect that you observed in the results and write one sentence on the interpretation of this results. For example, you can write about: Write which factor is associated with the outcome? Is there a significant interaction effect? Does anything change after adjusting for the covariate?

```{r}
# Preparation
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
dat$GLUCX <- as.numeric(as.character(dat$GLUCX))
dat$kw100.e <- as.numeric(as.character(dat$kw100.e))
dat$bmiKH <- factor(dat$bmiKH, labels = c("strong underweight", "underweight", "normal weight", "overweight", "obese"))
dat$sex <- factor(dat$sex, labels = c("boys", "girls"))
```

```{r}
#Prep
str(dat$GLUCX)
str(dat$kw100.e)
str(dat$sex)
str(dat$bmiKH)

# Look at distribution:
hist(dat$GLUCX)
# -> close to normal distribution
```
(i-ii) 1-way ANOVA model computation
```{r}
fit1_alt <- lm(GLUCX ~ bmiKH, data = dat)
summary(fit1_alt)
anova(fit1_alt)

#Both the F value and P value in the aforementioned example are small. Consequently, the glucose level changes little as BMI categories change.
```
```{r}
fit2_alt <- lm(GLUCX ~ sex, data = dat)
summary(fit2_alt)
anova(fit2_alt)

#Here the P value is really modest, yet the F value in this case is 91.71. Therefore, a significant change in glucose levels is observed as the sex groups change.

```
```{r}
# With anova function:
fit3.alt <- lm(GLUCX ~ bmiKH + sex, data = dat)
summary(fit3.alt)
anova(fit3.alt)

#The change in sex groups impacts the glucose level more considerably than the change in BMI categories, according to P and F values. bmiKH has a F value of 6.76, while sex has a F value of 89.68.
```
Model with both factors and their interaction
```{r}
# Visualization in profile plots
# install.packages("interactions")
library(interactions)

# Preparation: the function cat_plot for creating profile plots needs the results from aov() as input, computed with a dataset without missing values. i.e. first create a dataset without missing values:

data_for_analysis <- data.frame(GLUCX = dat$GLUCX, bmiKH = dat$bmiKH, sex = dat$sex)
data_for_analysis <- data_for_analysis[complete.cases(data_for_analysis),]
data_for_analysis <- droplevels(data_for_analysis)

# Now recompute ANOVA and use the cat_plot function.
fit4.neu <- aov(GLUCX ~ bmiKH + sex + bmiKH:sex, data = data_for_analysis)
cat_plot(fit4.neu, pred = bmiKH, modx = sex)

cat_plot(fit4.neu, pred = bmiKH, modx = sex, plot.points = TRUE)
(plot1 <- cat_plot(fit4.neu, pred = bmiKH, modx = sex, geom = "line") )

# -> The last line saves the plot in the object plot1 (so we may work with it later) and also displays the right-side findings.

# Compare to model without interaction:
cat_plot(aov(GLUCX ~ bmiKH + sex, data = data_for_analysis), pred = bmiKH, modx = sex, geom = "line")

#There are some interaction but they don't appear to be very substantial.Underweight and obesity is significantly differs in between both sex.
```
(v)
```{r}
data_for_analysis2 <- data.frame(GLUCX = dat$GLUCX, bmiKH = dat$bmiKH, sex = dat$sex, physical_wellbeing = dat$kw100.e)
data_for_analysis2 <- data_for_analysis2[complete.cases(data_for_analysis2),]
data_for_analysis2 <- droplevels(data_for_analysis2)

# Now recompute ANOVA and use the cat_plot function.
fit.neu <- aov(GLUCX ~ bmiKH + sex + physical_wellbeing + bmiKH:sex , data = data_for_analysis2)
cat_plot(fit.neu, pred = bmiKH, modx = sex)

cat_plot(fit.neu, pred = bmiKH, modx = sex, plot.points = TRUE)
(plot1 <- cat_plot(fit.neu, pred = bmiKH, modx = sex, geom = "line") )

#After adjusting the covariate we can see slight changes in the plot. For example: there is a change in underweight section for boys as it was slightly up of the GLUCX level 93, which now exactly lies at 93.
```



## Exercise 2: Posthoc tests in analysis of variance (optional)

Perform posthoc t-tests with and without a correction for multiple testing for the ANOVA in exercise 1, where we tested whether systolic blood pressure differed between age groups.

```{r}
# Preparation
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
dat$age2 <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
```

