---
title: "Homework 3"
author: "Shahriar Shohid Choudhury"
date: "November 7, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Compute frequencies in the Pima diabetes dataset

Load the Pima diabetes dataset:

```{r}
load(file = url("https://www.dropbox.com/s/hxn1c7yw9r8aw8v/Pima_diabetes.RData?dl=1"))
```

Which variables are measured on a nominal level?

```{r}
head(Pima_diabetes)
str(Pima_diabetes)
```
only the outcome (diabetes) nominal


Now compute frequency tables, barplots, and mosaic plots of all nominal variables in the dataset.
```{r}
# ann: 0 - does not have diabetes; 1 - has diabetes
Pima_diabetes$Diabetes <- factor(Pima_diabetes$Outcome, labels = c("no", "yes"))
# Absolute frequencies table:
table(Pima_diabetes$Outcome)
table(Pima_diabetes$Diabetes)
table_diab <- table(Pima_diabetes$Diabetes)

#relative frequencies table:

table(Pima_diabetes$Diabetes) / nrow(Pima_diabetes)

# barplot
barplot(table_diab, xlab = "Diabetes", ylab = "Absolute frequencies")

# mosaic plot
mosaicplot(table_diab, main = "Mosaic plot of Diabetes", color = TRUE)

```

Next, create a variable which describes whether a woman had more or less than 4 pregnancies. Then, use this variable to create a 2x2 table with diabetes outcome. Do you see an indication of whether the number of pregnancies is associated with diabetes prevalence? Do you think your investigation is a good way to investigate this?

```{r}
Pima_diabetes$Preg4 <- 0
Pima_diabetes$Preg4[Pima_diabetes$Pregnancies >= 4] <- 1
#select and assign women with 4 pregnancies to either group
Pima_diabetes$Preg4 <- factor(Pima_diabetes$Preg4, labels = c("< 4", ">= 4"))
table(Pima_diabetes$Preg4, Pima_diabetes$Diabetes, dnn = c("Number of pregnancies", "Diabetes"))

# I think the table seems not balanced. more women without diabetes >=4 pregnancies than women with diabetes and <4 pregnancies.
```

## Exercise 2: Generate a table with descriptive statistics (optional, but recommended)

Use any dataset (a dataset that you have worked with in the past, or that you are currently working with, a dataset that is available on Blackboard, in R or that you have downloaded from the internet), and generate a table with descriptive statistics of the main variables of interest.

## Exercise 3: Plots using ggplot2

Load the NoShow dataset:

```{r}
load(file = url("https://www.dropbox.com/s/4oqg79cn1qfnhsh/NoShowdata.RData?dl=1"))
head(NoShowdata)
```

Use ggplot2 to generate the following plots:

- Create a boxplots of Age (stratified) by neighborhood.
- Create a histogram of Age.
- Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.
- Stratify this plot further by gender.

What insights can you get from these plots? For which goal would you create these plots?

```{r}
#load ggplot2 package and format variables

library(ggplot2)
# corrected one negative value of age, and set it to NA
NoShowdata$Age[NoShowdata$Age < 0] <- NA
NoShowdata$Gender <- factor(NoShowdata$Gender, labels = c("female", "male"))
NoShowdata$Diabetes <- factor(NoShowdata$Diabetes, labels = c("FALSE", "TRUE"))
NoShowdata$Noshow <- factor(NoShowdata$"No-show", labels = c("did show up", "did not show up"))
NoShowdata$Neighbourhood <- factor(NoShowdata$Neighbourhood)

```


```{r}

# Boxplots
ggplot(data = NoShowdata, aes(y = Age)) + geom_boxplot()

```

```{r}
# -> Age distribution seems relative symmetric, however slightly skewed to higher age groups

ggplot(data = NoShowdata, aes(y = Age, x = Neighbourhood)) + geom_boxplot()

```

```{r}
# Histogram of age
ggplot(data = NoShowdata, mapping = aes(x = Age)) + geom_histogram()

```

```{r}
# Histogram of age stratified by NoShow
ggplot(data = NoShowdata, mapping = aes(x = Age, col = Noshow)) + geom_histogram()
ggplot(data = NoShowdata, mapping = aes(x = Age, fill = Noshow)) + geom_histogram()
ggplot(data = NoShowdata, mapping = aes(x = Age)) + geom_histogram() + facet_grid( ~ Noshow)
ggplot(data = NoShowdata, mapping = aes(x = Age)) + geom_histogram() + facet_grid(Gender ~ Noshow)

#There were many people that did show up for their appointment.

```

```{r}
# Histogram of age stratified by NoShow and Gender
ggplot(data = NoShowdata, mapping = aes(x = Age, fill = Noshow)) + geom_histogram() + facet_grid(~ Gender)
# The age distributions of those that didn't show up seems to be different between male and female. It is normally distributed for female but right skewed for males. More number of female showed up in compared to male counterpart.
```

From above we can see that younger people tend to miss more appointment.Between age group 15 to 60 missed more number of appointments.