---
title: "Homework 5"
author: "Shahriar Shohid Choudhury"
date: "November 20, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Probability distributions

Explore the functions *rnorm*, *rt*, *runif*, *rbinom* in R that allow you to generate random numbers from the normal, t-, uniform, and binomial distribution. Compute them with different values, and inspect histograms to visualize their distribution.

```{r}
# Binomial dist
# Generate data (1000 observations) from the normal distribution with three different expected values and variances:
X1a <- rbinom(n = 1000, size = 1, prob = 0.1)
X1b <- rbinom(n = 1000, size = 10, prob = 0.5)
X1c <- rbinom(n = 1000, size = 10, prob = 0.1)
#plotting histograms with the relative frequency on the Y axis
hist(X1a, freq = FALSE)
par(mfrow = c(1, 2))
hist(X1b, freq = FALSE, ylim = c(0, 1))
hist(X1c, freq = FALSE, ylim = c(0, 1))

```
```{r}
# Uniform dist

X2a <- runif(n = 1000, min = -1, max = 1)
X2b <- runif(n = 1000, min = 0, max = 2)
X2c <- runif(n = 1000, min = -5, max = 5)

#plotting histograms with the relative frequency on the Y axis
par(mfrow = c(1, 3))
hist(X2a, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 1))
hist(X2b, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 1))
hist(X2c, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 1))
```

```{r}
# Normal dist

X3a <- rnorm(n = 1000, mean = 0, sd = 1)
X3b <- rnorm(n = 1000, mean = 0, sd = 4)
X3c <- rnorm(n = 1000, mean = 4, sd = 0.5)

#plotting histograms with the relative frequency on the Y axis
par(mfrow = c(1, 3))
hist(X3a, freq = FALSE, xlim = c(-15, 15), ylim = c(0, 1))
hist(X3b, freq = FALSE, xlim = c(-15, 15), ylim = c(0, 1))
hist(X3c, freq = FALSE, xlim = c(-15, 15), ylim = c(0, 1))
```
```{r}
# t-dist
X4a <- rt(n = 1000, df = 1000)
X4b <- rt(n = 1000, df = 100)
X4c <- rt(n = 1000, df = 10)
X4d <- rt(n = 1000, df = 2)

#plotting histograms with the relative frequency on the Y axis
par(mfrow = c(1, 2))
hist(X3a, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.5))
hist(X4a, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.5))
par(mfrow = c(2, 2))
hist(X4a, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.5))
hist(X4b, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.5))
hist(X4c, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.5), breaks = 20)
hist(X4d, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.5), breaks = 100)
# -> the smaller df, the more extreme values there are

# the histogram shows the empirical distribution of the variable
par(mfrow = c(1, 1))
hist(X3a, freq = FALSE, col = "grey")
lines(density(X3a), col = "blue")

# To get the area under the curve (probability!) in the tails, the functions qnorm etc. can be used which computes the quantile:
qnorm(p = 0.95, mean = 0, sd = 1)
qnorm(p = 0.975, mean = 0, sd = 1)

```

## Exercise 2: Odds ratio

In the KiGGS dataset:

a) Compute the proportion of mothers that had hypertension during pregnancy. Use the variable 'e0155' which has values "Ja" (yes), "Nein" (No) and "Weiß nicht" (don't know).
b) Create a new variable that is 0 or 1 depending on whether the children are small or tall (think of a good way how to do this) based on the variable 'groeB' (body height). 
c) Then compute the odds ratio that the mother had hypertension during pregnancy (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children.

```{r}
# load data
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06

# a
str(dat$e0155)
table(dat$e0155)
# -> there are many missing values
#One possible estimate is

202/(2306+202)

# = 0.081. This would yield the estimate that 8.1% of mothers had hypertension during pregnancy.
```
```{r}
#b
# First, lets look at the variables:
str(dat$groeB)  # height
str(dat$age2)   # age (in groups of 2 years)
# Let's look at mean age by age group:
tapply(dat$groeB, dat$age2, mean, na.rm = TRUE)
# Now let's compute the new 0-1 tallness variable by checking for each child, whether it is taller or shorter than the average height of all children in their age group:
dat$tall <- NULL
for(i in 1:length(levels(dat$age2))){
  idx <- (dat$age2 == levels(dat$age2)[i])
  dat$tall[idx] <- ifelse(dat$groeB[idx] > mean(dat$groeB[idx], na.rm = TRUE), 1, 0)
}
dat$tall <- factor(dat$tall, labels = c("short", "tall"))
# reorder:
dat$tall <- factor(dat$tall, levels = c("tall", "short"))
table(dat$tall)
```
```{r}
# first, format the hypertension variable and remove the observations "Weiß nicht"

table(dat$e0155)
dat$e0155[dat$e0155 == "Weiß nicht"] <- NA
dat$e0155 <- droplevels(dat$e0155)
# check:
table(dat$e0155)

# compute cross table:
table(dat$e0155, dat$tall)

# proportion of (the mothers) having hypertension for tall kids:
p.tall <- 83/(83+857)
# proportion of (the mothers) having hypertension for short kids:
p.short <- 118/(118+1404)

# odds of (the mothers) having hypertension for tall kids:
odds.tall <- p.tall/(1-p.tall)
# odds of (the mothers) having hypertension for short kids:
odds.short <- p.short/(1-p.short)

# Odds ratio: 
odds.tall/odds.short

# Computation in short:
(83*1404)/(857*118)
```

## Exercise 3 (optional): Confidence intervals

Look at the hypertension variable from exercise 2. Use the binom::binom.confint and the questionr::odds.ratio functions to compute the estimates of the proportion and odds ratio as well as their confidence intervals. (you need to download and load these packages at first).

## Exercise 4 (optional, advanced): Bootstrap

Adapt the bootstrap implementation in R_5b_estimation_bootstrap.Rmd to compute the bootstrap estimate of the standard error of the variance of a normally-distributed and a t-distributed variable. Are they similar?