---
title: "Homework 7"
author: "Shahriar Shohid Choudhury"
date: "December 5, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.

```{r}
str(kiggs$KreaUrin)
table(!is.na(kiggs$KreaUrin))


str(kiggs$ERYT)
table(!is.na(kiggs$ERYT))

```
b) Compute the correlation between these two variables

#turning y as numeric value
```{r}

kiggs$ERYT <- as.numeric(as.character(kiggs$ERYT))

cor.test(kiggs$KreaUrin, kiggs$ERYT, use = "complete.obs", method = "pearson")

```
c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.

```{r}
library(mice)
library(VIM)
dat <- data.frame(KreaUrin= as.numeric(as.character(kiggs$KreaUrin)),
                  ERYT= as.numeric(as.character(kiggs$ERYT)))

tempData <- mice(dat, m = 5, maxit = 5, meth = 'pmm', seed = 50)
?mice
```
```{r}
summary(tempData)

#imputed Dataset
completedData1 <- complete(tempData,1)
completedData2 <- complete(tempData,2)
```


d) Compute the correlation again on the multiple imputed datasets.

```{r}
modelFit1 <- with(tempData, cor(KreaUrin, ERYT, method = "pearson"))
mean(unlist(modelFit1$analyses))
```
