---
title: "Homework 8"
author: "Shahriar Shohid Choudhury"
date: "December 11, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Load KiGGS data

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
```
## Exercise 1: Correlation

a) In the KiGGS dataset, compute the Pearson and Spearman correlation coefficient for the two variables 'sys1' and 'sys2' and hypothesis tests whether the two variables are associated or not. Interpret the results, and decide which of the two coefficients you would report in your analysis and why.

```{r}
# Two Metric Variable
sbp1 <- as.numeric(as.character(dat$sys1))
sbp2 <- as.numeric(as.character(dat$sys2))

# Correlations
cor(data.frame(sbp1, sbp2), use = "complete.obs", method = "pearson")
cor(data.frame(sbp1, sbp2), use = "complete.obs", method = "spearman")

#Hypothesis test:
cor.test(sbp1, sbp2, use = "complete.obs")

#For deciding which result to report we will look for histogram
hist(sbp1)
hist(sbp2)

# Here p values are small so enough evidence to reject the null hypothesis. I would report Pearson correlation because both are metric variables. And also normally distributed. On the other hand, Spearman correlation works better in ordinal values. 
```

b) Optional: Compute confidence intervals of the correlation coefficient estimates from part a). Note: for confidence intervals of the Spearman coefficient, you need another function.

## Exercise 2: Linear regression

a) Predict sys2 by sys1 using a simple linear regression, and interpret the results.

```{r}
# create variables
sys2 <- dat$sys2
sys1 <- dat$sys1

# Check format
str(sys2)
str(sys1)

# transform sys2 and sys1 to numeric variable!
sys2 <- as.numeric(as.character(sys2))
sys1 <- as.numeric(as.character(sys1))
```
```{r}

# compute linear regression models
res1 <- lm(sys2 ~ sys1)

# Look at the results
summary(res1)

#R^2 =0.72 that means Sys1 explains 72% of the variation of sys2. 
#The p value is low, indicating that there is a link between sys1 and sys2.
```
```{r}

# Visualize results
plot(sys1, sys2)
abline(a = summary(res1)$coefficients[1, 1], b = summary(res1)$coefficients[2, 1])

# get predictions of Sys2
sys2_pred1 <- predict(res1)
head(sys2)
head(sys2_pred1)
```
b) Add age2 and sex as predictors to the linear regression model above, and interpret the results. 

```{r}
#new variables for age2 and sex 
x <- dat$age2
y <- dat$sex

# check format
str(x)
str(y)
table(x)
table(y)


#factorization
x <- factor(x, labels = c("0 - 1 J.", "2 - 3 J.", "4 - 5 J.", "6 - 7 J.", "8 - 9 J.", "10 - 11 J.", "12 - 13 J.", "14 - 15 J.", "16 - 17 J."))
y <- factor(y, labels = c("Männlich", "Weiblich"))

# check again
table(x)
table(y)

#Data are treated correctly by modeling functions when they are stored as factors.

```
```{r}
# New linear regression model
res2 <- lm(sys2 ~ sys1+ as.numeric(x)+as.numeric(y))
# Look at the results
summary(res2)

#R^2=0.74
```
```{r}
# Prediction
sys2_pred2 <- predict(res2)
head(sys2)
head(sys2_pred2)

#R^2=0.74, Age and sex do not significantly contribute to the variance of sys2 beyond what is already explained by sys1, but this does not imply that they do not contribute significantly to the variance at all.
```

## Exercise 3: Visualization of regression (optional)

Use the functions in ggplot2 to compute a scatter plot and insert the regression line of the analysis in exercise 2a.