---
title: "Final exam"
author: "Shahriar Shohid Choudhury-815468"
date: "February 26, 2023"
output:
  html_document: default
  pdf_document: default
---

#Question 02

## (a) Import KiGGS data

```{r}
library(haven)
r <- haven::read_sav("KiGGS03_06.sav")
```

## (b) New dataframe

```{r}
kiggs <- r[, c("E070M", "E070V", "E072", "E074", "arztZ01", "sex", "age2", "schichtz", "e065z")]
kiggs
```

## (c) Format and save the dataframe

```{r}
# Dataframe saving
save(kiggs, file = "kiggs.Rdata")
```

##Question 03

## (a)

#variables factor check
```{r}
#Checking the variables E070M, E070V, E072, E074 are all factors or not
is.factor(kiggs$E070M)
is.factor(kiggs$E070V)
is.factor(kiggs$E072)
is.factor(kiggs$E074)
#All the mentioned variables are factor
```
#value changing
```{r}
#Changing all children's values for the variable E074's "has not breastfed" value to NA
levels(kiggs$E074)[levels(kiggs$E074)=="has not breastfed"]<-NA
```

#deleting empty factor
```{r}
kiggs <- kiggs[!is.na(kiggs$E074), ]
```

#checking the steps worked or not
```{r}
table(is.na(kiggs$E074))
```
#calculating new variable
```{r}
library(dplyr)
kiggs$burdenS <- dense_rank(kiggs$E070M)+ dense_rank(kiggs$E070V)+ dense_rank(kiggs$E072)+dense_rank(kiggs$E074)
table(kiggs$burdenS)
#A high value indicates that the children were exposed to low level of smoking. As the rank increased 6599 were exposed with the low level of smoking. In contrast, 134 were exposed to the highest level of smoking.
```
## (b) Add new variable

#updating
```{r}
save(kiggs, file = "kiggs.Rdata")
```


##Question 04

## (a) Descriptive statistics calculation
```{r,echo = FALSE, results = "asis"}
kiggs$arztZ01 <- as.numeric(as.character(kiggs$arztZ01))
round_val <- 3
```

```{r,echo = FALSE, results = "asis"}
library(qwraps2)
options(qwraps2_markup = "markdown")

r1 <-
  list(
    "arztZ01" = list(
      "mean" = ~ round(mean(arztZ01, na.rm = TRUE), digits = round_val),
      "sd" = ~ round(sd(arztZ01, na.rm = TRUE), digits = round_val)),
    "sex" = list(
      "boys" = ~ qwraps2::n_perc0(sex == "boys", show_symbol = TRUE),
      "girls" = ~ qwraps2::n_perc0(sex == "girls", show_symbol = TRUE)),
    "age2" = list(
      "0-1y" = ~ qwraps2::n_perc0(age2 == "0-1y", show_symbol = TRUE),
      "2-3y" = ~ qwraps2::n_perc0(age2 == "2-3y", show_symbol = TRUE),
      "4-5y" = ~ qwraps2::n_perc0(age2 == "4-5y", show_symbol = TRUE),
      "6-7y" = ~ qwraps2::n_perc0(age2 == "6-7y", show_symbol = TRUE),
      "8-9y" = ~ qwraps2::n_perc0(age2 == "8-9y", show_symbol = TRUE),
      "10-11y" = ~ qwraps2::n_perc0(age2 == "10-11y", show_symbol = TRUE),
      "12-13y" = ~ qwraps2::n_perc0(age2 == "12-13y", show_symbol = TRUE),
      "14-15y" = ~ qwraps2::n_perc0(age2 == "14-15y", show_symbol = TRUE),
      "16-17y" = ~ qwraps2::n_perc0(age2 == "16-17y", show_symbol = TRUE)),
    "burdenS" = list(
      "4" = ~ qwraps2::n_perc0(burdenS == "4", show_symbol = TRUE, na_rm = TRUE),
      "5" = ~ qwraps2::n_perc0(burdenS == "5", show_symbol = TRUE, na_rm = TRUE),
      "6" = ~ qwraps2::n_perc0(burdenS == "6", show_symbol = TRUE, na_rm = TRUE),
      "7" = ~ qwraps2::n_perc0(burdenS == "7", show_symbol = TRUE, na_rm = TRUE),
      "8" = ~ qwraps2::n_perc0(burdenS == "8", show_symbol = TRUE, na_rm = TRUE),
      "9" = ~ qwraps2::n_perc0(burdenS == "9", show_symbol = TRUE, na_rm = TRUE),
      "10" = ~ qwraps2::n_perc0(burdenS == "10", show_symbol = TRUE, na_rm = TRUE),
      "11" = ~ qwraps2::n_perc0(burdenS == "11", show_symbol = TRUE, na_rm = TRUE),              
      "12" = ~ qwraps2::n_perc0(burdenS == "12", show_symbol = TRUE, na_rm = TRUE))
  )
    qwraps2::summary_table(kiggs, r1)
    #The mean for number of paediatrician visits is 2.464 and the sd is 3.395
```
##(b) Missing values

- arztZ01 has `r sum(is.na(kiggs$arztZ01))` missing values
- sex has `r sum(is.na(kiggs$sex))` missing values
- age2 has `r sum(is.na(kiggs$age2))` missing values
- burdenS has `r sum(is.na(kiggs$burdenS))` missing values
- `r sum(complete.cases(kiggs))` observations have complete data

##Question 05

## (a)Calculate the linear regression

- arztZ01 is already a metric variable which is being used as outcome.
- As sex, age2, schichtz and e065z are factor variables which need to      converted into metric variables as they are being used as predictor.

```{r}
# Computation of the regression model
model_lm <- lm(arztZ01 ~ burdenS + as.numeric(sex) + as.numeric(age2) + as.numeric(schichtz) + as.numeric(e065z), data = kiggs)
```

## (b) Significance of coefficients
```{r}




```